\documentclass[12pt]{article}  % Default 10pt → 12pt body font

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{ragged2e}

\makeatletter
\renewcommand*{\maketitle}{%
  \vspace*{2cm}
  { \raggedleft  % Scoped: Header only
    {\LARGE\bfseries Problem Set 1 \par}
    \vspace{1em}
    {\Large Brent Myers \par}
    \vspace{1em}
    {\large 01/27/26 \par}
  }
  \vspace{2em}
}
\makeatother



\begin{document}
\maketitle

\section{Introduction}

Twenty years ago, I graduated with a degree in economics from Southern Utah University and began a successful 15-year career in financial services with Fidelity Investments in Salt Lake City, Utah. I left Fidelity to attend Utah Valley University, where I earned a master's degree in accountancy before transitioning into public accounting as an international tax and transfer pricing specialist. This experience of returning to graduate school and pivoting to international tax and transfer pricing ignited my desire to pursue a PhD in accounting, ultimately leading me to the University of Oklahoma.

As a first-year PhD student in accounting, I am taking this class to learn how to handle data efficiently and effectively. Although my professional background involved extensive data use for strategic decision-making, it included very little coding or data wrangling. Therefore, I am excited, but somewhat nervous, to learn the core concepts of data science and gain foundational skills in the field.

For my class project, I plan to investigate the key determinants of income shifting by multinational enterprises, which involves transferring taxable income to lower-tax jurisdictions to reduce the overall tax burden. Data for this topic are sparse and often messy, making this class particularly valuable for developing my ability to work with such challenging datasets.

After graduation, I hope to join a research university where I can conduct studies in transfer pricing and international taxation while teaching courses in these areas.

\section{Equation}

\begin{equation}
a^2 + b^2 = c^2 \label{eq:pythagoras}
\end{equation}

\end{document}
